Sperm should evolve to make female meiosis fair.

### 2014-05-18 writeup_and_notes/male_dependent_drive.tex
added hypotheses to discussion

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozygote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
	provide evolution  with time and opportunity to further minimize drive.
This is because such drive suppressing alleles reduced their chances
of forming low fitness homozygotes. 
%In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
%	consistent with the tenuous evidence that the non-driving
%        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
%        sperm \cite{agulnick, Pomiankowski1993}.   
% YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
More generally, alleles that act through sperm to
 reduce the opportunity of female meiotic drive regardless of linkage or phase.

@@

The interests of fertilizing sperm and female are quite well aligned during syngamy -- producing a viable and potentially fit offspring. 
%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meiosis. 
We predict that female drive should be

\yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
	(1) MII drivers should drive less often when fertilized by complimentary sperm. 
	(2) Drive may be more severs in interspecific crosses in which sperm have not specifically evolved to resists a driver. 
	(3) ??? Any more specific hypotheses -- maybe making this last sentence more explicit, e.g. self-fertilizing or otherwise frequently inbreeding animals will not require fertilization for the completion of meiosis... [note - this runs counter to our empirical examples of mice and C-elegens]}

>>>>

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozygote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
	provide evolution  with time and opportunity to further minimize drive.
This is because such drive suppressing alleles reduced their chances
of forming low fitness homozygotes. 
%In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
%	consistent with the tenuous evidence that the non-driving
%        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
%        sperm \cite{agulnick, Pomiankowski1993}.   
% YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
More generally, alleles that act through sperm to
 reduce the opportunity of female meiotic drive regardless of linkage or phase. 
We predict that female drive, at a stage influencable by meiotic
drive,  should be reduced when fertilized by
complimentary sperm. For example, MII female drive should be more severe in
interspecific crosses as sperm will not specifically evolved to
resists a female driver.

@@

The interests of fertilizing sperm and female are quite well aligned during syngamy -- producing a viable and potentially fit offspring. 
%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
	only species' different reproductive ecologies, but also
        different selection pressures shaping the interaction of sperm
        alleles with female meiosis. 
For example it would be interesting to know whether changes that are predicted to lower the opportunity for female meiotic drive,
        e.g. an evolutionary transition to inbreeding or a high
        selfing rate in animals, are often accompanied by a relaxation of the requirement of fertilization for the completion of meiosis.

% \yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
% 	(1) MII drivers should drive less often when fertilized by complimentary sperm. 
% 	(2) Drive may be more severs in interspecific crosses in which sperm have not specifically evolved to resists a driver. 
% 	(3) ??? Any more specific hypotheses -- maybe making this last sentence more explicit, e.g. self-fertilizing or otherwise frequently inbreeding animals will not require fertilization for the completion of meiosis... [note - this runs counter to our empirical examples of mice and C-elegens]}
